Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-22 sections/eval.tex
Update on Overleaf.

There is a gap between the expressiveness of the purely declarative Array API and the Finch language which defines both outputs and the algorithm used to produce those outputs. The naive approach to solving this would be to explicitly define a mapping from each function in the API to a Finch program which eagerly produces the output. This approach requires a simple translation procedure but misses crucial opportunities for optimization across a chain of function calls and needlessly materializes large intermediate results. For example,

Due to the wide variety of formats and protocols supported by Finch, it allows for a large array of potential impleme

>>>>

There is a gap between the expressiveness of the purely declarative Array API and the Finch language which defines both the output and the algorithm. The naive approach to solving this would be to explicitly define a mapping from each function in the API to a Finch program which eagerly produces the output. This approach results in a simple translation procedure but misses crucial opportunities for optimization across a chain of function calls and needlessly materializes large intermediate results. To avoid this, we have implemented a lazy evaluation approach that collects API calls and then computes output when requested by the user. This is implemented through 1) Finch Logic, a minimal, high-level language for expressing array operations and 2) the Finch Interpreter, which lowers Finch Logic programs to one or more Finch programs while making heuristic decisions about output format, loop ordering, and protocols.
